---
title: "Data Visualization"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,warning=FALSE, message=FALSE}
library (dplyr)
library (tidyr)
library (ggplot2)
library (scales)
library (leaflet)
library (DT)
# to call stats filter - stats::filter()
```


for dyplr/tidyr reference see my [data tidying page](https://yiyi063.github.io/arctic_training/data_formating)

# reformat the data for plotting

Alaska Department of Fish and Game, Division of Commercial Fisheries. 2017. Daily salmon escapement counts from the OceanAK database, Alaska, 1921-2017. Knowledge Network for Biocomplexity. doi:10.5063/F1S46Q6M. 

```{r}
esc <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e", method = "libcurl"),
                    stringsAsFactors = FALSE)

```

*calculate annual escapment by species and region
+'seperate'
+'group by'
+'summarise' using the 'sum' function

```{r}
annual_esc <- esc %>% 
  separate(sampleDate, c("Year", "Month", "Day"), sep = "-") %>% 
    #coerce year to numeric type
 mutate(Year=as.numeric(Year)) %>% 
  
  group_by(Species, SASAP.Region, Year) %>% 
  summarize(count = sum(DailyCount)) %>% 
  filter(Species %in% c("Chinook", "Sockeye", "Chum", "Coho", "Pink"))

head(annual_esc)
```


species_mean <-  catch_clean %>% 
  group_by(species) %>% 
  summarise(catch_mean = mean(catch),num_obs = n())


*filter for just pacific salmon (Sockeye,Goho,Pink, Chun)


#generate some static plots

```{r}
ggplot(annual_esc,aes(x=Species, y=count))+
  geom_col()
```

```{r}
ggplot(annual_esc,mapping=aes(x=Species, y=count))+
  geom_col(fill="Blue")
```
```{r}
ggplot(annual_esc, aes(x = Species, y = count, fill = SASAP.Region)) +
  geom_col()
```

```{r}
ggplot(filter(annual_esc,SASAP.Region=="Kodiak"),aes(x=Year,y=count,color=Species))+
  geom_line()+
  geom_point()+
  ylab("Escapement") +
    ggtitle("Kodiak Salmon Escapement") +
    theme_test()+
  theme(legend.position="bottom",legend.title=element_blank())
```

```{r}
p <- ggplot(annual_esc, aes(x = Year, y = count, color = Species)) + 
    geom_line() +
    geom_point() +
    facet_wrap(~SASAP.Region, scales = "free_y") +
    ylab("Escapement") +
    scale_y_continuous(labels = comma) +
    theme(legend.position="bottom",legend.title=element_blank())

ggsave("figures/regional_escapment.png", 
       plot=p,
       width=7,
       height=8,
       units="in")
  
```


#generate an interactive map

map the sampling location for escapment 

* find distinct lat/lon/location combination

```{r}
locations <- esc %>% 
  distinct(Location, Latitude, Longitude) %>% 
  drop_na()
```

```{r,echo=F}
datatable(locations) %>% 

```
```{r}
leaflet(locations) %>% 
  addTiles() %>% 
 addMarkers(~ Longitude, ~ Latitude, popup = ~ Location)
```


