---
title: "Data formating"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,warning=FALSE, message=FALSE}
library (dplyr)
library (tidyr)

# to call stats filter - stats::filter()
```


# Reformat Catch Data


* remove "all" columns
* create species column
  - move from wide to long
* general QA

```{r}
catch_original <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", method = "libcurl"),
                    stringsAsFactors = FALSE)
head(catch_original)
```


Remove the "all" and "notesRegCode" columns using "select"

Control + shift + m: piple operator shortcut: %>%

```{r}
catch_long <- catch_original %>% 
  select(-All, -notesRegCode) %>% 
  gather(key="species", value="catch", -Region, -Year)

head(catch_long)
```
```{r}
catch_wide <- catch_long %>% 
  spread(key=Region, value=catch)

head(catch_wide)
```
## clean up our data

* rename catch to catch_thousands
* change catch column to numeric
* create a new catch column in units num. of fish

```{r}
catch_clean <- catch_long %>% 
  rename(catch_thousands = catch) %>%
  mutate (catch_thousands = ifelse(catch_thousands == "I",1,catch_thousands)) %>%  
  mutate(catch_thousands=as.numeric(catch_thousands)) %>% 
  mutate(catch=catch_thousands *1000) %>% 
  select(-catch_thousands)

head (catch_clean)
```


```{r, eval=FALSE,echo=FALSE}
#look for where as.numeric failed
test <- as.numeric(catch_long$catch)
i <- which(is.na(test)==TRUE)
catch_long[i,]
```
## split - Apply - Combine

* calcualte mean catch by species
```{r}
species_mean <-  catch_clean %>% 
  group_by(species) %>% 
  summarise(catch_mean = mean(catch),num_obs = n())

head(species_mean)
```

* calcualte mean catch by species and year
```{r}
year_mean <- catch_clean %>% 
  group_by(Year, species) %>% 
  summarise(catch_mean=mean(catch)) %>% 
  arrange(-catch_mean) %>% 
  filter(Year >= 1990 | Year <=1900)

year_mean
```

# Join the Region Definitions

```{r}
region_defs <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1",
                            method = "libcurl"),
                        stringsAsFactors = FALSE) %>% 
    select(code, mgmtArea)

head(region_defs)
```
```{r}
catch_joined <- left_join(catch_clean,region_defs, by=c("Region"="code"))

head(catch_joined)

```


# Misc. Functions

seperate and unite

```{r}
dates_df <- data.frame(date = c("5/24/1930",
                                "5/25/1930",
                                "5/26/1930",
                                "5/27/1930",
                                "5/28/1930"),
                       stringsAsFactors = FALSE)

dates_df %>% 
  separate(date, c("month", "day", "year"), "/")
```
```{r}
dates_split <- dates_df %>% 
  separate(date, into=c("month","day","year"),sep='/',remove=F)

head(dates_split)
```

```{r}
dates_unite <- dates_split %>% 
  unite(col= date_iso,year,month,day,sep="-")

head(dates_unite)
```

```{r}
library(stringr)
str_pad("5",2,side='left',pad='0')

```

```{r}
write.csv(catch_clean,"catch_clean.csv",row.names = F)
```

